Demo arXiv template

### 2022-10-26 paper/paper.qmd
changes after meeting + a subsection on a toy example

---
title: "Demo arXiv template"
format:
  arxiv-pdf:
    keep-tex: true  
  arxiv-html: default
author:
  - name: H. Sherry Zhang
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, VIC
        country: Australia
    orcid: 0000-0002-7122-1463
    email: huize.zhang@monash.edu
    url: https://huizezhangsh.netlify.app/
  - name: Collaborators
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, VIC
        country: Australia
abstract: |
  - Indices, useful, quantify severity, early monitoring, 
  - A huge number of indices have been proposed by domain experts, however, a large majority of them are not being adopted, reused, and compared in research or in practice. 
  - One of the reasons for this is the plenty of indices are quite complex and there is no obvious easy-to-use implementation to apply them to user's data. 
  - The paper describes a general pipeline framework to construct indices from spatio-temporal data, 
  - This allows all the indices to be constructed through a uniform data pipeline and different indices to vary on the details of each step in the data pipeline and their orders. 
  - The pipeline proposed aim to smooth the workflow of index construction through breaking down the complicated steps proposed by various indices into small building blocks shared by most of the indices.
  - The framework will be demonstrated with drought indices as examples, but appliable in general to environmental indices constructed from multivariate spatio-temporal data
keywords: 
  - spatio-temporal data
  - indices
  - data pipeline
bibliography: bibliography.bib  
---

@@

## Tidy data {#sec-tidy-data}

As a scripted language for data analysis, R provides a interactive environment for users to code with data without the abstraction a typical programming language requires. 
This flexible environment is paramount for data analysis since in practice, data analysis is an iterative process that involves trials and errors to solve the problems.
Analysts need the freedom to quickly experiment with different ideas when working with data.
The tidyverse suite provides a collection of R package for conducting data analysis.
These tools directly operate on data, and users, even new users, can easily read from the code on what operation is performed on the data and what to expect as the output.
All of these are built upon a simple prior knowledge of the tidy data principle [@wickham_tidy_2014], which prescribes three rules to structure tabular data: 1) Each variable forms a column, 2) Each observation forms a row, and; 3) Each type of observational unit forms a table.

@@

<!-- While data can be arranged in various different ways, maintaining a fixed representation gives a predictable output of operations on data.  -->

<!-- Only when each step in the workflow has been mapped out, do data analysts need to consider those abstract concept of object oriented programming and functional programming to help them design the workflow.  -->
<!-- how raw data should be processed to the final results can only be consolidated when  -->

<!-- Before the concept of tidy data [@wickham_tidy_2014], tabular data arrive at data analysts in all different ways.  -->
<!-- Different analysts would write customised scripts for analysing the specific data.  -->
<!-- These scripts can be extended to other data analysed by the same people or group but this is not generalizable directly to another dataset.  -->
<!-- When the tidy data concept comes, variables and values are arranged so that 1) Each variable forms a column, 2) Each observation forms a row, and; 3) Each type of observational unit forms a table. -->
<!-- With this specific layout, wrangling on tabular data can be standardised into a grammar of data manipulation in `dplyr` [@dplyr].  -->

<!-- A similar issue happens with index construction where researchers construct their own indices in their own ways and there has not yet been a tidy principle on index construction. -->
<!-- Also, this tidy principle on index construction is more complex than those in tidy data and the `dplyr` package. -->
<!-- It has to encompass the workflow of transformation from the raw data towards the final index series. -->

## Data pipeline {#sec-data-pipeline}

The concept of pipeline (and pipe) is not new concept to computing. It refers to a set of data processing elements connected in series, where the output of one element is the input of the next one.
@wickham_plumbing_2009 argues that whether made explicit or not, pipeline has to be presented in every graphics program.
The paper also argues that breaking down graphic rendering into steps is beneficial for understand the implementation and compare between different graphic systems. 
Both arguments are applicable to index construction.
Papers that proposing new indices always present a workflow diagram to illustrate how data is transformed into the final indices, but not much implementation has explicitly present this this step-by-step transformation in the code design.

To better understand the pipeline approach in statistic computing, examples in interactive graphics and machine learning models are given.
In interactive graphics, the data pipeline is used to describe how the data is processed to be rendered on the screen. 
Examples of early work on building pipeline for interactive graphics includes @buja_elements_1988, @sutherland_orca_2000, and @xie_reactive_2014. 
Steps in these pipelines include non-linear transformation, variable standardization, randomization and dimension reduction.
A more familiar example from tidymodels [@tidymodels] presents a set of packages for machine learning models following the tidyverse principles [@tidyverse]. 
The machine learning pipeline constitutes data resampling, feature engineering, model fitting, model tuning, and model evaluation each being responsible by a separate package. 
Before tidymodel, analysts who wish to fit different models need to learn the workflow of each individual algorithm and need to rewrite most of the model fitting, tuning, and evaluation code when switching from one model to another. 
The tidymodel approach unifies a wide range of models under a single pipeline where users can substitute different models while maintain the same pipeline.
This allows users to easily experiment with different machine learning models a lot easier.

@@

@@

## Raw data

**Another section on original data directly downloaded, can have different spatial resolution, temporal granularity, data quality problem. After processing them and align them together they become the "raw data"**

@@

## Spatial aggregation

@@

## Scaling

@@

## Normalising

@@

## Variable transformation

@@

## Temporal processing

## Dimension reduction

@@

## Benchmarking

## Simplification

>>>>

---
title: "Demo arXiv template"
format:
  arxiv-pdf:
    keep-tex: true  
  arxiv-html: default
author:
  - name: H. Sherry Zhang
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, VIC
        country: Australia
    orcid: 0000-0002-7122-1463
    email: huize.zhang@monash.edu
    url: https://huizezhangsh.netlify.app/
  - name: Collaborators
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, VIC
        country: Australia
abstract: |
  - Indices, useful, quantify severity, early monitoring, 
  - A huge number of indices have been proposed by domain experts, however, a large majority of them are not being adopted, reused, and compared in research or in practice. 
  - One of the reasons for this is the plenty of indices are quite complex and there is no obvious easy-to-use implementation to apply them to user's data. 
  - The paper describes a general pipeline framework to construct indices from spatio-temporal data, 
  - This allows all the indices to be constructed through a uniform data pipeline and different indices to vary on the details of each step in the data pipeline and their orders. 
  - The pipeline proposed aim to smooth the workflow of index construction through breaking down the complicated steps proposed by various indices into small building blocks shared by most of the indices.
  - The framework will be demonstrated with drought indices as examples, but appliable in general to environmental indices constructed from multivariate spatio-temporal data
keywords: 
  - spatio-temporal data
  - indices
  - data pipeline
bibliography: bibliography.bib  
editor_options: 
  chunk_output_type: inline
---

@@

The concept of pipeline (and pipe) is not a new concept to computing. It refers to a set of data processing elements connected in series, where the output of one element is the input of the next one.
@wickham_plumbing_2009 argues that whether made explicit or not, pipeline has to be presented in every graphics program.
The paper also argues that breaking down graphic rendering into steps is beneficial for understand the implementation and compare between different graphic systems.

*The pipeline system would work partially because of an agreed data structure (tidy data) and [same input and output format] - why the pipeline approach would work.*

The tidyverse suite provides a collection of R package for conducting data analysis.
These tools directly operate on data, and users, even new users, can easily read from the code on what operation is performed on the data and what to expect as the output.
All of these are built upon a simple prior knowledge of the tidy data principle [@wickham_tidy_2014], which prescribes three rules to structure tabular data: 1) Each variable forms a column, 2) Each observation forms a row, and; 3) Each type of observational unit forms a table.

@@

To better understand the pipeline approach in statistic computing, examples in interactive graphics and machine learning models are given.
In interactive graphics, the data pipeline is used to describe how the data is processed to be rendered on the screen. 
Examples of early work on building pipeline for interactive graphics includes @buja_elements_1988, @sutherland_orca_2000, and @xie_reactive_2014. 
Steps in these pipelines include non-linear transformation, variable standardization, randomization and dimension reduction.

A more familiar example from tidymodels [@tidymodels] presents a set of packages for machine learning models following the tidyverse principles [@tidyverse]. 
The machine learning pipeline constitutes data resampling, feature engineering, model fitting, model tuning, and model evaluation each being responsible by a separate package. 
Before tidymodel, analysts who wish to fit different models need to learn the workflow of each individual algorithm and need to rewrite most of the model fitting, tuning, and evaluation code when switching from one model to another. 
The tidymodel approach unifies a wide range of models under a single pipeline where users can substitute different models while maintain the same pipeline.
This allows users to easily experiment with different machine learning models a lot easier.

Both arguments are applicable to index construction.
Papers that proposing new indices always present a workflow diagram to illustrate how data is transformed into the final indices, but not much implementation has explicitly present this this step-by-step transformation in the code design.

@@

## How does the pipeline constructin of an index look like?

Consider a commonly used drought index: Standardised Precipitation Index (SPI) [@spi]. Its construction involves the following steps: 1) aggregate monthly summed precipitation into a time scale $k$, 2) fit the aggregated series into a distribution to get the density values, and 3) convert the density values to normal quantiles. Using the pipeline approach, the index will be constructed as:

```{r eval = FALSE, echo = TRUE}
DATA %>% 
  ts_aggregate(scale = 12, col = prcp) %>% 
  normalising(dist = "gamma", col = prcp, fit_method = "lmom") %>% 
  augment(col = "prcp_agg")
```

*The exact code needs to be further turned: 1) use verb: ts_aggreate -> aggregate_ts?, normalising -> normalise, 2) using `var` instead of `col` since it may be confused with "color", 3) normalising -> fit? since technically this step is no longer to normalise the fitted density value into normal quantiles but a distribution fitting step.*

Here the column `prcp` is first temporally *aggregated* with a time scale of 12. The output is then normalised with a gamma distribution, fitted using the L-moment. The `augment` step, following the verb in the `broom` package, accepts the fitted distributions and adds information about each month in the dataset to get the final index.

```{r eval = FALSE, echo = FALSE}
res <- w %>% 
  ts_aggregate(scale = 12, date = ym, id = id, col = prcp) %>% 
  normalising(dist = "gamma", date = ym, id = id, col = prcp,
              method = "lmom") %>% 
  augment(id = id, date = ym,  col = "prcp_agg")

res <- res %>% 
  rename(
         pipeline_d = prcp_agg,
         pipeline_idx = .index)

data(wichita)
wichita <- ts(wichita[,-c(1,2)], end=c(2011,10), frequency=12)
spi12 <- spi(wichita[,'PRCP'], 12)
from_SPI <- tibble(
  ym = yearmonth(zoo::yearmon(zoo::index(wichita))), 
  SPI_idx = as.vector(spi12$fitted)
)

compare <- res %>% left_join(from_SPI, by = "ym") 
compare %>% 
  ggplot(aes(x = SPI_idx, y = pipeline_idx)) + 
  geom_point() + 
  theme_bw() + 
  theme(legend.position = "bottom", 
        aspect.ratio = 1) +
  ggtitle("Index values of Standardised Precipitation Index (SPI)", 
          subtitle = "using wichita data from the SPEI package") + 
  xlab("Calculated from a single command") + 
  ylab("Calculated from pipeline")

```

```{r fig-toy-example, eval = TRUE, echo = FALSE}
#| fig-align: center
#| fig-cap: "this is the caption"
#| out-height: 50%
#| out-width: 50%
knitr::include_graphics(here::here("figures/toy-example-spi.png"))
```

While the calculated index value from the pipeline approach is the same as those calculated using a single line command like `spi(DATA, ARGUMENTS)`, as shown in @fig-toy-example, the pipeline approach gives more flexibility to perform diagnostics on the intermediate steps and to extend on the initial index calculation. For example, model goodness of fit test can be checked after `normalising()` through:

```{r eval = FALSE, echo = TRUE}
normalising(dist = "gamma", col = d, fit_method = "lmom") %>% 
  gof_test()
```

and bootstrap uncertainty on the index, with 100 boostrap samples, can be calculated using `n_boot = 100` in the `normalising()` as:

```{r eval = FALSE, echo = TRUE}
normalising(dist = "gamma", col = d, fit_method = "lmom", n_boot = 100) 
```

## Pipeline steps for constructing indies

*put a diagram here to summarise the pipeline*

@@

each step as a point rather than subsection

**Raw data**

*Another section on original data directly downloaded, can have different spatial resolution, temporal granularity, data quality problem. After processing them and align them together they become the "raw data"*

@@

**Spatial aggregation**

@@

**Scaling**

@@

**Normalising**

@@

**Variable transformation**

@@

**Temporal processing**

**Dimension reduction**

@@

**Benchmarking**

**Simplification**

### 2022-10-26 paper/paper.tex
changes after meeting + a subsection on a toy example

@@

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, sharp corners, breakable, borderline west={3pt}{0pt}{shadecolor}, frame hidden, enhanced, interior hidden]}{\end{tcolorbox}}\fi

@@

\hypertarget{sec-tidy-data}{%
\subsection{Tidy data}\label{sec-tidy-data}}

As a scripted language for data analysis, R provides a interactive
environment for users to code with data without the abstraction a
typical programming language requires. This flexible environment is
paramount for data analysis since in practice, data analysis is an
iterative process that involves trials and errors to solve the problems.
Analysts need the freedom to quickly experiment with different ideas
when working with data. The tidyverse suite provides a collection of R
package for conducting data analysis. These tools directly operate on
data, and users, even new users, can easily read from the code on what
operation is performed on the data and what to expect as the output. All
of these are built upon a simple prior knowledge of the tidy data
principle (Wickham 2014), which prescribes three rules to structure
tabular data: 1) Each variable forms a column, 2) Each observation forms
a row, and; 3) Each type of observational unit forms a table.

@@

\hypertarget{sec-data-pipeline}{%
\subsection{Data pipeline}\label{sec-data-pipeline}}

The concept of pipeline (and pipe) is not new concept to computing. It
refers to a set of data processing elements connected in series, where
the output of one element is the input of the next one. Wickham et al.
(2009) argues that whether made explicit or not, pipeline has to be
presented in every graphics program. The paper also argues that breaking
down graphic rendering into steps is beneficial for understand the
implementation and compare between different graphic systems. Both
arguments are applicable to index construction. Papers that proposing
new indices always present a workflow diagram to illustrate how data is
transformed into the final indices, but not much implementation has
explicitly present this this step-by-step transformation in the code
design.

To better understand the pipeline approach in statistic computing,
examples in interactive graphics and machine learning models are given.
In interactive graphics, the data pipeline is used to describe how the
data is processed to be rendered on the screen. Examples of early work
on building pipeline for interactive graphics includes Buja et al.
(1988), Sutherland et al. (2000), and Xie, Hofmann, and Cheng (2014).
Steps in these pipelines include non-linear transformation, variable
standardization, randomization and dimension reduction. A more familiar
example from tidymodels (Kuhn and Wickham 2020) presents a set of
packages for machine learning models following the tidyverse principles
(Wickham et al. 2019). The machine learning pipeline constitutes data
resampling, feature engineering, model fitting, model tuning, and model
evaluation each being responsible by a separate package. Before
tidymodel, analysts who wish to fit different models need to learn the
workflow of each individual algorithm and need to rewrite most of the
model fitting, tuning, and evaluation code when switching from one model
to another. The tidymodel approach unifies a wide range of models under
a single pipeline where users can substitute different models while
maintain the same pipeline. This allows users to easily experiment with
different machine learning models a lot easier.

@@

@@

\hypertarget{raw-data}{%
\subsection{Raw data}\label{raw-data}}

\textbf{Another section on original data directly downloaded, can have
different spatial resolution, temporal granularity, data quality
problem. After processing them and align them together they become the
``raw data''}

@@

\hypertarget{spatial-aggregation}{%
\subsection{Spatial aggregation}\label{spatial-aggregation}}

@@

\hypertarget{scaling}{%
\subsection{Scaling}\label{scaling}}

@@

\hypertarget{normalising}{%
\subsection{Normalising}\label{normalising}}

@@

\hypertarget{variable-transformation}{%
\subsection{Variable transformation}\label{variable-transformation}}

@@

\hypertarget{temporal-processing}{%
\subsection{Temporal processing}\label{temporal-processing}}

\hypertarget{dimension-reduction}{%
\subsection{Dimension reduction}\label{dimension-reduction}}

@@

\hypertarget{benchmarking}{%
\subsection{Benchmarking}\label{benchmarking}}

\hypertarget{simplification}{%
\subsection{Simplification}\label{simplification}}

@@

>>>>

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

@@

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, sharp corners, frame hidden, breakable, borderline west={3pt}{0pt}{shadecolor}, interior hidden, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

The concept of pipeline (and pipe) is not a new concept to computing. It
refers to a set of data processing elements connected in series, where
the output of one element is the input of the next one. Wickham et al.
(2009) argues that whether made explicit or not, pipeline has to be
presented in every graphics program. The paper also argues that breaking
down graphic rendering into steps is beneficial for understand the
implementation and compare between different graphic systems.

\emph{The pipeline system would work partially because of an agreed data
structure (tidy data) and {[}same input and output format{]} - why the
pipeline approach would work.}

The tidyverse suite provides a collection of R package for conducting
data analysis. These tools directly operate on data, and users, even new
users, can easily read from the code on what operation is performed on
the data and what to expect as the output. All of these are built upon a
simple prior knowledge of the tidy data principle (Wickham 2014), which
prescribes three rules to structure tabular data: 1) Each variable forms
a column, 2) Each observation forms a row, and; 3) Each type of
observational unit forms a table.

@@

To better understand the pipeline approach in statistic computing,
examples in interactive graphics and machine learning models are given.
In interactive graphics, the data pipeline is used to describe how the
data is processed to be rendered on the screen. Examples of early work
on building pipeline for interactive graphics includes Buja et al.
(1988), Sutherland et al. (2000), and Xie, Hofmann, and Cheng (2014).
Steps in these pipelines include non-linear transformation, variable
standardization, randomization and dimension reduction.

A more familiar example from tidymodels (Kuhn and Wickham 2020) presents
a set of packages for machine learning models following the tidyverse
principles (Wickham et al. 2019). The machine learning pipeline
constitutes data resampling, feature engineering, model fitting, model
tuning, and model evaluation each being responsible by a separate
package. Before tidymodel, analysts who wish to fit different models
need to learn the workflow of each individual algorithm and need to
rewrite most of the model fitting, tuning, and evaluation code when
switching from one model to another. The tidymodel approach unifies a
wide range of models under a single pipeline where users can substitute
different models while maintain the same pipeline. This allows users to
easily experiment with different machine learning models a lot easier.

Both arguments are applicable to index construction. Papers that
proposing new indices always present a workflow diagram to illustrate
how data is transformed into the final indices, but not much
implementation has explicitly present this this step-by-step
transformation in the code design.

@@

\hypertarget{how-does-the-pipeline-constructin-of-an-index-look-like}{%
\subsection{How does the pipeline constructin of an index look
like?}\label{how-does-the-pipeline-constructin-of-an-index-look-like}}

Consider a commonly used drought index: Standardised Precipitation Index
(SPI) (McKee, Doesken, and Kleist 1993). Its construction involves the
following steps: 1) aggregate monthly summed precipitation into a time
scale \(k\), 2) fit the aggregated series into a distribution to get the
density values, and 3) convert the density values to normal quantiles.
Using the pipeline approach, the index will be constructed as:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{DATA }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{ts\_aggregate}\NormalTok{(}\AttributeTok{scale =} \DecValTok{12}\NormalTok{, }\AttributeTok{col =}\NormalTok{ prcp) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{normalising}\NormalTok{(}\AttributeTok{dist =} \StringTok{"gamma"}\NormalTok{, }\AttributeTok{col =}\NormalTok{ prcp, }\AttributeTok{fit\_method =} \StringTok{"lmom"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{augment}\NormalTok{(}\AttributeTok{col =} \StringTok{"prcp\_agg"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\emph{The exact code needs to be further turned: 1) use verb:
ts\_aggreate -\textgreater{} aggregate\_ts?, normalising -\textgreater{}
normalise, 2) using \texttt{var} instead of \texttt{col} since it may be
confused with ``color'', 3) normalising -\textgreater{} fit? since
technically this step is no longer to normalise the fitted density value
into normal quantiles but a distribution fitting step.}

Here the column \texttt{prcp} is first temporally \emph{aggregated} with
a time scale of 12. The output is then normalised with a gamma
distribution, fitted using the L-moment. The \texttt{augment} step,
following the verb in the \texttt{broom} package, accepts the fitted
distributions and adds information about each month in the dataset to
get the final index.

\begin{figure}

{\centering \includegraphics[width=0.5\textwidth,height=0.5\textheight]{../figures/toy-example-spi.png}

}

\caption{\label{fig-toy-example}this is the caption}

\end{figure}

While the calculated index value from the pipeline approach is the same
as those calculated using a single line command like
\texttt{spi(DATA,\ ARGUMENTS)}, as shown in
Figure~\ref{fig-toy-example}, the pipeline approach gives more
flexibility to perform diagnostics on the intermediate steps and to
extend on the initial index calculation. For example, model goodness of
fit test can be checked after \texttt{normalising()} through:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{normalising}\NormalTok{(}\AttributeTok{dist =} \StringTok{"gamma"}\NormalTok{, }\AttributeTok{col =}\NormalTok{ d, }\AttributeTok{fit\_method =} \StringTok{"lmom"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{gof\_test}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

and bootstrap uncertainty on the index, with 100 boostrap samples, can
be calculated using \texttt{n\_boot\ =\ 100} in the
\texttt{normalising()} as:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{normalising}\NormalTok{(}\AttributeTok{dist =} \StringTok{"gamma"}\NormalTok{, }\AttributeTok{col =}\NormalTok{ d, }\AttributeTok{fit\_method =} \StringTok{"lmom"}\NormalTok{, }\AttributeTok{n\_boot =} \DecValTok{100}\NormalTok{) }
\end{Highlighting}
\end{Shaded}

\hypertarget{pipeline-steps-for-constructing-indies}{%
\subsection{Pipeline steps for constructing
indies}\label{pipeline-steps-for-constructing-indies}}

\emph{put a diagram here to summarise the pipeline}

@@

each step as a point rather than subsection

\textbf{Raw data}

\emph{Another section on original data directly downloaded, can have
different spatial resolution, temporal granularity, data quality
problem. After processing them and align them together they become the
``raw data''}

@@

\textbf{Spatial aggregation}

@@

\textbf{Scaling}

@@

\textbf{Normalising}

@@

\textbf{Variable transformation}

@@

\textbf{Temporal processing}

\textbf{Dimension reduction}

@@

\textbf{Benchmarking}

\textbf{Simplification}

@@

\leavevmode\vadjust pre{\hypertarget{ref-spi}{}}%
McKee, Thomas B, Nolan J Doesken, and John Kleist. 1993. {``The
Relationship of Drought Frequency and Duration to Time Scales.''} In
\emph{Proceedings of the 8th Conference of Applied Climatology},
179--84. Anaheim, CA: American Meteorological Society.
